---
title: "Homework 3"
author: Ben Lebwohl
output: github_document
---

## Problem 1

Load the instacart dataset
```{r}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
data("instacart")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%")
```

_Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations._

```{r}
instacart
view(instacart)
```

The data set is a tibble, with >1.3 million observations and 14 columns. Each row is an order of a product. These come in batches of orders (order_id) and are tied to a user (user_id). Each product is associated with an aisle (aisle_id) and department (department_id)

Let's see some of the more common departments and aisles that are ordered:

```{r}
instacart %>% 
  group_by(department, aisle) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  arrange(desc(n_obs))
```

_How many aisles are there, and which aisles are the most items ordered from?_

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  arrange(desc(n_obs))
```

There are 134 aisles, and the aisles that most items are ordered from are fresh vegetables and fresh fruits.

_Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it._

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  filter(n_obs >= 10000) %>% 
     mutate(aisle = fct_reorder(aisle, n_obs)) %>% 
  ggplot(aes(x = aisle, y = n_obs, label = aisle)) +
geom_point() +
scale_y_log10(breaks = c(20000, 40000, 60000, 80000, 100000, 150000), labels = c("20000", "40000", "60000", "80000", "100000", "150000")) +
  geom_text(aes(label=aisle, hjust = 0, vjust = 1), size = 3, check_overlap = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  labs(
    y = "Number of orders" 
  )
```


_Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table._
```{r}
instacart %>% 
  group_by(aisle, product_name) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  mutate(
    order_rank = min_rank(-n_obs)
  ) %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  filter(order_rank <= 3) %>% 
  arrange(aisle, order_rank)
```


_Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)._

```{r}
instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(names_from = product_name, values_from = mean_hour)
```

## Problem 2


_Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc)._

```{r}
accel_df =
read_csv("./data/accel_data.csv") %>% 
janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "minute", 
    names_prefix = "activity_", 
    values_to = "activity") %>% 
  mutate(
    type_of_day = ifelse(day == "Saturday" | day == "Sunday" , "weekend", "weekday")
  )

accel_df %>% 
  group_by(day_id) %>% 
  summarize(n_obs = n())
```

The data set contains are 35 days, each with activity data on each of the 1440 minutes of the day.

_Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?_

```{r}
accel_df %>% 
  group_by(day_id, type_of_day) %>% 
  summarize (total_activity = sum(activity)) %>% 
  ggplot(aes(x = day_id, y = total_activity, color = type_of_day)) +
  geom_point()
```

There is wide variability in activity! Weekends tend to have lower total activity.

_Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph._

```{r}
accel_df %>% 
  mutate(
    minute = as.numeric(minute)
    )%>% 
  ggplot(aes(x = minute, y = activity, color = day)) +
  geom_point() +
   scale_x_continuous(
    breaks = c(200, 400, 600, 800, 1000, 1200),
    labels = c("200", "400", "600", "800", "1000", "1200")
   )
```

There are discrete times of day in which activity is consistently increased.

## Problem 3

Load the NY NOAA data:

```{r}
library(p8105.datasets)
data("ny_noaa")
ny_noaa
```

This is a tibble with 2.59 million rows, containing data on precipitation and temperature for various weather stations

```{r}
ny_noaa %>% 
  group_by(id) %>% 
  summarize(n_obs = n())

ny_noaa %>% 
  summary()
```

There appear to be ~1500 weather stations, with dates spanning from 1981 through 2010.

_Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?_

```{r}
ny_noaa =
ny_noaa %>% 
  mutate(
  year = lubridate::year(date),
  month = lubridate::month(date),
  day = lubridate::day(date),
  tmin = as.numeric(tmin) /10,
  tmax = as.numeric(tmax) /10
  ) %>% 
  relocate(id, year, month, day) %>% 
  select(-date)

```

```{r}
ny_noaa %>% 
    group_by(snow) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  mutate(
    order_rank = min_rank(-n_obs)
    )%>% 
      arrange(order_rank)
```

For snowfall, the most commonly observed values are 0 and NA.

_Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?_

```{r}
  ny_noaa %>% 
  filter(month == 1 | month == 7) %>% 
  group_by(id, month) %>% 
  summarize(
    average_tmax = median(tmax, na.rm = TRUE)
  ) %>% 
  filter(average_tmax != "NaN") %>% 
ggplot(aes(x = average_tmax)) +
        geom_histogram() +
        facet_grid(. ~ month)
       
```

_Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year._

```{r}
 plot_1 = ny_noaa %>% 
  ggplot(aes (x = tmin, y = tmax)) +
           geom_hex()

plot_2 = ny_noaa %>% 
  filter(snow <100 & snow >0) %>% 
  mutate(
    year = as.factor(year)
  ) %>% 
  ggplot(aes(y = snow)) +
  geom_boxplot() +
    facet_grid(. ~ year) +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot_1 / plot_2
```

